Plots

### 2025-02-25 src/plots.Rmd
Working func with patchwork

```{r, echo=FALSE, eval=FALSE}
install.packages("tidyverse")
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
```

@@

@@

@@

@@

# Calc num of records for each age
n_all <- df %>% group_by(age) %>% count
n_adult <- n_all %>% filter(age == "adult") %>% pull(n)

@@

# Calc pccc
df_plot <- df_long %>%
    group_by(age_name, cod_name, model_name, .drop = F) %>%
    summarize( # calc pccc
        pccc = calc_pccc(cod, model_cghr10, N = unique(ncod)[1])
    ) %>%
    mutate( # add pccc to model name
        model_label = paste0(model_name, " (", round(pccc, 2), ")")
    ) %>%
    filter(age_name == "Adult")

# Add min and max values for each cause of death
df_ends <- df_plot %>%
    group_by(cod_name) %>%
    summarize(
        min_pccc = min(pccc),
        max_pccc = max(pccc),
        model_min = model_name[which.min(pccc)],
        model_max = model_name[which.max(pccc)]
    ) %>%
    arrange(max_pccc)

# Sort cod by max model pccc
df_plot <- df_plot %>%
    mutate(cod_name = factor(cod_name, levels = df_ends$cod_name))
df_ends <- df_ends %>%
    mutate(cod_name = factor(cod_name, levels = df_ends$cod_name))

# Filter out the mid models
df_mids <- df_plot %>%
    left_join(df_ends, by = "cod_name") %>%
    filter(model_name != model_max & model_name != model_min)

# Plot
ggplot(df_plot, aes(x = pccc, y = cod_name)) +
    geom_point( # mid points
        data = df_mids,
        color = "gray60",
        size = 1
    ) +
    geom_point (# min points
        data = df_ends,
        aes(x = min_pccc),
        size = 1.5
    ) +
    geom_point (# max points
        data = df_ends,
        aes(x = max_pccc),
        size = 1.5
    ) +
    geom_segment( # min max lines
        data = df_ends,
        aes(
            x = min_pccc,
            xend = max_pccc,
            y = cod_name,
            yend = cod_name
        ),
        linewidth = 0.75
    ) +
    geom_text( # min model text
        data = df_ends,
        aes(
            x = min_pccc,
            y = cod_name,
            label = paste0(model_min, " (", round(min_pccc, 2), ")")
        ),
        hjust = 1.2,
        vjust = 0.5,
        size = 3
    ) +
    geom_text( # max model text
        data = df_ends,
        aes(
            x = max_pccc,
            y = cod_name,
            label = paste0(model_max, " (", round(max_pccc, 2), ")")
        ),
        hjust = -0.2,
        vjust = 0.5,
        size = 3
    ) +
    geom_text_repel( # mid model text
        data = df_mids,
        aes(label = model_label),
        color = "gray60",
        max.overlaps = Inf,
        size = 2
    ) +
    scale_x_continuous(expand = expansion(add = 0.2)) +
    labs(
        x = "Partial Chance Corrected Concordance (PCCC)",
        y = paste0(length(unique(df_plot$cod_name)), " Causes of Death for ", n_adult, " Adult Records")
    ) +
    theme_minimal() +
    theme(
        panel.border = element_rect(
            color = "black",
            fill = NA,
            linewidth = 0.5
        ),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )

>>>>

```{r, echo=FALSE, eval=FALSE}
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggrepel")
#install.packages("patchwork")
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(patchwork)
```

@@

## Metric Functions

@@

## Plot Functions

Functions for generating plots.

# Plot performance of models
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        title = NULL,
        x_title = NULL,
        y_title = NULL,
        text_mid_size = 2,
        text_end_size = 3,
        text_hjust = 0.2,
        x_limits = c(0, 1),
        ...
    ) {

    # Calc metric
    df_plot <- df_long %>%
        group_by(.data[[group]], model_name, .drop = F) %>%
        summarize( # calc pccc
            metric = f(cod, model_cghr10, N = unique(ncod)[1], ...)
        ) %>%
        mutate( # add pccc to model name
            model_label = paste0(model_name, " (", round(metric, 2), ")")
        )

    # Add min and max values for each cause of death
    df_ends <- df_plot %>%
        group_by(.data[[group]]) %>%
        summarize(
            min_metric = min(metric),
            max_metric = max(metric),
            model_min = model_name[which.min(metric)],
            model_max = model_name[which.max(metric)]
        ) %>%
        arrange(max_metric)

    # Sort cod by max model metric
    df_plot[[group]] <- factor(df_plot[[group]], levels = df_ends[[group]])
    df_ends[[group]] <- factor(df_ends[[group]], levels = df_ends[[group]])

    # Filter out the mid models
    df_mids <- df_plot %>%
        left_join(df_ends, by = group) %>%
        filter(model_name != model_max & model_name != model_min)

    # Plot metric for each model
    out <- ggplot(df_plot, aes(x = metric, y = .data[[group]])) +
        geom_point( # mid points
            data = df_mids,
            color = "gray60",
            size = 1
        ) +
        geom_point (# min points
            data = df_ends,
            aes(x = min_metric),
            size = 1.5
        ) +
        geom_point (# max points
            data = df_ends,
            aes(x = max_metric),
            size = 1.5
        ) +
        geom_segment( # min max lines
            data = df_ends,
            aes(
                x = min_metric,
                xend = max_metric,
                y = .data[[group]],
                yend = .data[[group]]
            ),
            linewidth = 0.75
        ) +
        geom_text( # min model text
            data = df_ends,
            aes(
                x = min_metric,
                y = .data[[group]],
                label = paste0(model_min, " (", round(min_metric, 2), ")")
            ),
            hjust = 1 + text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text( # max model text
            data = df_ends,
            aes(
                x = max_metric,
                y = .data[[group]],
                label = paste0(model_max, " (", round(max_metric, 2), ")")
            ),
            hjust = 0 - text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text_repel( # mid model text
            data = df_mids,
            aes(label = model_label),
            color = "gray60",
            max.overlaps = Inf,
            size = text_mid_size
        ) +
        scale_x_continuous(
            limits = x_limits,
            expand = expansion(add = 0.2)
        ) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            panel.border = element_rect(
                color = "black",
                fill = NA,
                linewidth = 0.5
            ),
            legend.position = "none",
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank()
        )
    return(out)
}

@@

## Shorten COD Names

Shorten long names into new lines.

# Shorten COD names
df <- df %>%
    mutate(physician_cghr10 = str_wrap(physician_cghr10, width = 34))

@@

# Calc num of records for each age
n_all <- df %>% group_by(age) %>% count
n_adult <- n_all %>% filter(age == "adult") %>% pull(n)
n_child <- n_all %>% filter(age == "child") %>% pull(n)
n_neo <- n_all %>% filter(age == "neo") %>% pull(n)

# Calc num of cods for each age
cod_all <- df_long %>% group_by(age) %>% summarize(n = n_distinct(cod))
cod_adult <- cod_all %>% filter(age == "adult") %>% pull(n)
cod_child <- cod_all %>% filter(age == "child") %>% pull(n)
cod_neo <- cod_all %>% filter(age == "neo") %>% pull(n)

@@

# Plot perf for pccc
plot_perf(
    df_long %>% filter(age == "adult"),
    calc_pccc,
    title = paste0(cod_adult, " Causes of Death for ", n_adult, " Adult Records")
) /
    plot_perf(
        df_long %>% filter(age == "child"),
        calc_pccc,
        title = paste0(cod_child, " Causes of Death for ", n_child, " Child Records")
    ) /
    plot_perf(
        df_long %>% filter(age == "neo"),
        calc_pccc,
        title = paste0(cod_neo, " Causes of Death for ", n_neo, " Neonatal Records")
    ) +
    plot_layout(heights = c(2, 1, 1))
